\textit{Supplementary materials for: } High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2021-06-04 paper/main.tex
minor formatting edits

Overall, we found that we could reliably recover both first-order and
second-order correlations from the synthetic data
(Fig.~\ref{fig:higher_order_sims}).  When the correlations were stable
for longer intervals, or changed gradually (constant, ramping, and
event datasets), recovery performance was relatively high, and we were
better able to recover dynamic first-order correlations than
second-order correlations.  This is because errors in our \textit{estimation}
procedure at lower orders necessarily propagate to higher orders
(since lower-order correlations are used to estimate higher-order
correlations).  Conversely, when the correlations were particularly
\textit{un}stable (random datasets), we better recovered second-order
correlations.  This is because noise in our \textit{data generation}
procedure propagates from higher orders to lower orders (see
\textit{Synthetic data: simulating dynamic high-order correlations}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first-order and second-order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic higher-order
      correlations}.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  \textbf{Constant.}  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  \textbf{Random.} These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint.  \textbf{Ramping.}
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints.  \textbf{Event.}  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

We also examined the impact of the data duration (Fig.~\timeheatmap)
and complexity (number of zero-order features; Fig.~\features) on our
ability to accurately recover ground truth first-order and
second-order dynamic correlations.  In general, we found that our approach
better recovers ground truth dynamic correlations from longer duration
timeseries data.

We also found that our approach tends to best recover data generated
using fewer zero-order features (i.e., lower complexity), although
this tendency was not strictly monotonic.  Further, because our data
generation procedure requires $\mathcal{O}(K^{4})$ memory to generate
a second-order timeseries with $K$ zero-order features, we were not
able to fully explore how the number of zero-order features affects
recovery accuracy as the number of features gets larger (e.g., as it
approaches the number of features present in the fMRI data we examine
below).  Although we were not able to formally test this to our
satisfaction, we expect that accurately estimating dynamic high-order
correlations would require data with many more zero-order features
than we were able to simulate.  Our reasoning is that
high-order correlations necessarily involve larger numbers of
lower-order features, so achieving adequate ``resolution'' high-order
timeseries might require many low-order features.

>>>>

Overall, we found that we could reliably recover both first-order and
second-order correlations from the synthetic data
(Fig.~\ref{fig:higher_order_sims}).  When the correlations were stable
for longer intervals, or changed gradually (constant, ramping, and
event datasets), recovery performance was relatively high, and we were
better able to recover dynamic first-order correlations than
second-order correlations.  This is because errors in our \textit{estimation}
procedure at lower orders necessarily propagate to higher orders
(since lower-order correlations are used to estimate higher-order
correlations).  Conversely, when the correlations were particularly
\textit{un}stable (random datasets), we better recovered second-order
correlations.  This is because noises in our \textit{data generation}
procedure propagates from higher orders to lower orders (see
\textit{Synthetic data: simulating dynamic high-order correlations}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first-order and second-order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic higher-order
      correlations}.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  \textbf{Constant.}  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  \textbf{Random.} These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint.  \textbf{Ramping.}
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints.  \textbf{Event.}  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.  Note that the ``dips''
    and ``ramps'' at the boundaries of sharp transitions (e.g., the
    beginning and ends of the ``constant'' and ``ramping'' datasets,
    and at the event boundaries of the ``event'' datasets) are
    finite-sample effects that reflect the reduced numbers of samples
    that may be used to accurately estimate correlations at sharp
    boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

We also examined the impact of the data duration (Fig.~\timeheatmap)
and complexity (number of zero-order features; Fig.~\features) on our
ability to accurately recover ground truth first-order and
second-order dynamic correlations.  In general, we found that our approach
better recovers ground truth dynamic correlations from longer duration
timeseries data.  We also found that our approach tends to best recover data generated
using fewer zero-order features (i.e., lower complexity), although
this tendency was not strictly monotonic.  Further, because our data
generation procedure requires $\mathcal{O}(K^{4})$ memory to generate
a second-order timeseries with $K$ zero-order features, we were not
able to fully explore how the number of zero-order features affects
recovery accuracy as the number of features gets larger (e.g., as it
approaches the number of features present in the fMRI data we examine
below).  Although we were not able to formally test this to our
satisfaction, we expect that accurately estimating dynamic high-order
correlations would require data with many more zero-order features
than we were able to simulate.  Our reasoning is that
high-order correlations necessarily involve larger numbers of
lower-order features, so achieving adequate ``resolution'' high-order
timeseries might require many low-order features.

### 2021-06-04 paper/supplementary_materials.tex
updates to supp and main text

@@

\newpage
\renewcommand{\refname}{Supplementary references}
\bibliography{CDL-bibliography/memlab}

>>>>

\begin{figure}[p!]
\centering
\includegraphics[width=0.95\textwidth]{figs/sim_heatmaps_time}
  \caption{\textbf{Recovery heatmap of simulated first-order and
      second-order dynamic correlations across time and widths.}  Each panel displays a
    heatmap of the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices.
    The averages are taken across several randomly generated
    synthetic datasets for each timeseries pattern (constant, random,
    ramping, and data; see \textit{Synthetic data: simulating dynamic higher-order
      correlations}).  The $x$-axes of each heatmap denote varying
    Laplace-shaped kernel widths (see Figs.~\ref{fig:supp_kernels}
    and~\ref{fig:supp_widths}), and the $y$-axes of each heatmap
    denote varying durations (in samples) of the synthetic datasets.
    A total of 10 synthetic datasets were generated for each duration
    (row) and timeseries pattern (panel).}
\label{fig:time}
\end{figure}

\begin{figure}[p!]
\centering
\includegraphics[width=0.95\textwidth]{figs/sim_heatmaps_features}
  \caption{\textbf{Recovery heatmap of simulated first-order and
      second-order dynamic correlations across features and widths.}  Each panel displays a
    heatmap of the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices.
    The averages are taken across several randomly generated
    synthetic datasets for each timeseries pattern (constant, random,
    ramping, and data; see \textit{Synthetic data: simulating dynamic higher-order
      correlations}).  The $x$-axes of each heatmap denote varying
    Laplace-shaped kernel widths (see Figs.~\ref{fig:supp_kernels}
    and~\ref{fig:supp_widths}), and the $y$-axes of each heatmap
    denote varying numbers of features ($K$) used in each of the synthetic datasets.
    A total of 10 synthetic datasets were generated for each number of features
    (row) and timeseries pattern (panel).}
\label{fig:features}
\end{figure}

@@

\newpage
\renewcommand{\refname}{Supplementary references}
\bibliography{CDL-bibliography/cdl}
